---
title: "WUP_04"
author: "lele!"
date: "fall 2025"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats) 
GSScsv.csv <-read_csv("GSScsv.csv")
``` 


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
This data is numerical categorical and i will be using a helper file to help me solve the assignment.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(GSScsv.csv,aes(y=WORDSUM,x=DEGREE))
```
```{r}
basicCN + geom_boxplot()
```
```

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```
```{r}

favstats(WORDSUM~DEGREE,data=GSScsv.csv)
```

The graduate school Q1 is higher than the Q3 of  both the high school and LT high school. Suggesting there will be a statistical difference between them.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSScsv.csv)
summary.aov(model1)
```
When observing, I see the p-value is 0.000236. Which tells me the is a large diffenc between the categories

```{r}
par(mar=c(3,17,2,2))
plot(TukeyHSD(model1,conf.level=0.98),las=2,)
```

After plotting i see that 98% confidence level, the confidence4 doesn't pass through are LT high school Graduate and high School graduate. This sugest that they are significantly different.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->